Appendix for: Deep Reinforcement Learning for Conservation Decisions

### 2021-04-21 manuscript/appendix.Rmd
appendix

abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib
output: 
  rticles::arxiv_article:
    keep_tex: true
---

@@

```{r}
## reproducible settings
np <- import("numpy")
seed <- 24L # integer
np$random$seed(seed)
# Optionally set to "" to force CPU-evaluation if needing perfect reproducibility
# Sys.setenv("CUDA_VISIBLE_DEVICES"="")

@@

Standard R methods are used to plot the results summarized across the replicates:

@@

\begin{equation}
\mu_{t} = X_t + X_t r \left(1 - \frac{X_t}{K}) - \frac{a_t X_t^q}{X_t^q + b^q} 
X_{t+1} \sim \lognormal(\mut_t, \sigma)
\end{equation}

@@

\beign{equation}
a_{t+1} = a_t + \alpha - A_t
\end{equation}

@@

$$U(X_t, A_t) = b X_t - c A_t ^2$$
where we will take $b = 1$ and $c=1$.  
Our implementation in `gym_conservation` allows the user to consider alternative parameter choices, alternative models for the ecological dynamics, and alternate types of actions, such as manipulating the ecosystem state directly rather than manipulating the environmental parameter. 
For some such scenarios the optimal solution is known or can be determined by stochastic dynamic programming, while for others, including the scenario of focus here, the optimal solution is unknown.

@@

As discussed in the main text, we have no existing optimal solution to the tipping point problem, and so rely on a common heuristic strategy instead:
select a fixed level of conservation investment that is sufficient to counter-balance any further side towards the tipping point, preserving it in it's current state.

@@

@@

@@

@@

```{r}
# Some individual replicates, for comparison
sims_df %>%
  filter(rep < 100, time < 500) %>%
  ggplot(aes(time, state, col=model, group=interaction(model, rep))) +
  geom_line(alpha = .1, show.legend = FALSE) + facet_wrap(~model)
```

```{r}
# Some individual replicates, for comparison
is_low <- sims_df %>% filter(model == "TD3") %>% group_by(rep, model) %>% summarize(low = sum(state < .2) > 3) %>% filter(low)
sims_df %>% inner_join(is_low) %>%
  ggplot(aes(time, state,  group=interaction(model, rep))) +
  geom_line(color = pal[3], show.legend = FALSE) + facet_wrap(~rep)
```

```{r}
# Some individual replicates, for comparison
is_low <- sims_df %>% filter(model == "steady-state") %>% group_by(rep, model) %>% summarize(low = sum(state < .2) > 3) %>% filter(low) %>% head(16)
sims_df %>% inner_join(is_low) %>%
  ggplot(aes(time, state, group=interaction(model, rep))) +
  geom_line(color = pal[1], show.legend = FALSE) + facet_wrap(~rep)
```

@@

```{r fig2, fig.cap=""}
fig_sims / ( fig_policy + fig_reward)
```

>>>>

keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
bibliography: references.bib

header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}

output: 
  rticles::arxiv_article:
    keep_tex: true
---

@@

```{r}
## reproducible settings
np <- import("numpy")
seed <- 24L # integer
np$random$seed(seed)
# Optionally set to "" to force CPU-evaluation if needing perfect reproducibility
Sys.setenv("CUDA_VISIBLE_DEVICES"="")
set.seed(seed)
```

@@

<!-- Remaining is just plotting code, repeated in the main .Rmd.  Should we omit it from the appendix? -->
Standard R methods are used to plot the results summarized across the replicates:

@@

\[
\mu_{t} = X_t + X_t r \left(1 - \frac{X_t}{K} \right) - \frac{a_t X_t^q}{X_t^q + b^q}
\]
\[
X_{t+1} \sim \text{lognormal}(\mu_t, \sigma)
\]

@@

\[
a_{t+1} = a_t + \alpha - A_t
\]

@@

\[U(X_t, A_t) = b X_t - c A_t ^2\]

where we will take \(b = 1\) and \(c=1\).  
Our implementation in `gym_conservation` allows the user to consider alternative parameter choices, alternative models for the ecological dynamics, and alternate types of actions, such as manipulating the ecosystem state directly rather than manipulating the environmental parameter. 
For some such scenarios the optimal solution is known or can be determined by stochastic dynamic programming, while for others, including the scenario of focus here, the optimal solution is unknown.

@@

As discussed in the main text, we have no existing optimal solution to the tipping point problem, and so rely on a common heuristic strategy instead:
select a fixed level of conservation investment that is sufficient to counter-balance any further side towards the tipping point, preserving it in it's current state.
This is implemented using the `fixed_action` method provided in our `gym_conservation` module, which also implements other heursitic models, including a human agent which requires interactive input to select the action each year.

@@

We gather together the results under the RL agent and steady-state policy as before,

@@

The resulting three data frames contain the necessary data for each of the subplots in figure 3 of the main text.

@@

<!-- Remaining is just plotting code, repeated in the main .Rmd.  Should we omit it from the appendix? -->

@@

@@

```{r fig2, include=FALSE}
fig_sims / ( fig_policy + fig_reward)
```

Because it is difficult to get a feel for the dynamics of individual replicate simulations from ensemble statistics, we select a few example trajectories to examine directly. 
Of the 100 replicate simulations, we pick 4 examples that dip below a state value of 0.2 for over 15 consecutive timesteps, indicating a transition into the lower basin of attraction. 
Comparing the dynamics under the rule of thumb steady-state strategy to that of the RL-trained agent, it is clear that RL agent does a better job at both avoiding tipping points and promoting the recovery of those selected trajectories that cross into the lower attractor.

```{r}
# Some individual replicates, for comparison

## First 4 of the TD3 reps falling below .2 for more than 15 steps
is_low <- sims_df %>% 
  filter(model == "TD3") %>% 
  group_by(rep, model) %>% 
  summarize(low = sum(state < .2) > 10) %>% 
  filter(low) %>% head(6)

## First 4 such cases:
sims_df %>% inner_join(is_low) %>%
  ggplot(aes(time, state,  group=interaction(model, rep))) +
  geom_line(color = pal[3], show.legend = FALSE) + facet_wrap(~rep)
```

```{r}
# Some individual replicates, for comparison
is_low <- sims_df %>% filter(model == "steady-state") %>%
  group_by(rep, model) %>% summarize(low = sum(state < .2) > 10) %>% 
  filter(low) %>% head(6)
sims_df %>% inner_join(is_low) %>%
  ggplot(aes(time, state, group=interaction(model, rep))) +
  geom_line(color = pal[1], show.legend = FALSE) + facet_wrap(~rep)
```

### 2021-04-21 manuscript/cover-letter.Rmd
appendix

We are pleased to present a proposal for an article discussing the potential use of deep Reinforcement Learning in theoretical and applied ecological conservation.

>>>>

We are pleased to present a proposal for an article discussing the potential use of deep Reinforcement Learning in theoretical and applied ecological conservation for the Methods section of Ecology Letters.
